Getting Started with the Graded Response Model (GRM): A gentle introduction and tutorial in R

### 2024-01-16 draft_paper.qmd
Step 5 done!

To scrutinize their measurement quality, psychological researchers have relied upon two foundational principles underlying psychological testing, i.e., CTT and IRT, to decompose observed score into its *deterministic* and *random* or stochastic elements [@zumboValidityFoundationalIssues2006]. On the one hand, CTT decomposes observed score into *true* score and its *error* components. In practice, researchers implement this principle by conducting Structural Equation Modeling (SEM), Exploratory Factor Analysis (EFA), or Confirmatory Factor Analysis (CFA). Specifying the measurement part of SEM, or employing EFA and CFA models, allows for the decomposition of the covariance matrix of a given data into latent factors, which represent true score or actual ability or trait level, and error terms (i.e., residuals or unique variances). Consequently, the relationship between observed score and true score is *linear*, as true score is essentially a linear transformation of observed score [@embretsonItemResponseTheory2000]. IRT, on the other hand, models this relationship in a *probabilistic* rather than a linear fashion. More sharply, an IRT model assumes a probabilistic relationship between the observed score and the latent trait being measured ($\Theta$), and in particular, this probability takes into account the properties of the items, such as item discrimination, difficulty, and, if necessary, guessing [@bakerBasicsItemResponse2017].

@@

IRT offers an elegant solution to this issue by enabling the calculation of standard errors for each individual (*SE*~$\theta$~), and test reliability can be inferred from the average of *SE*~$\theta$~ across a sample of participants [@embretsonItemResponseTheory2000; @langSciencePracticeItem2021]. This approach allows for the estimation of reliability at varying levels of trait or ability (i.e., test information function - TIF). By doing this, when a test is overly challenging for low-performing individuals or too easy for high-performing individuals, IRT analysis can help identify the levels of trait or ability where the test is most reliable. A concrete example of this is a study scrutinizing the reliability of the Short Dark Tetrad (SD4) scale from an IRT perspective [@blotnerDarkTriadDead2022]. According to this study, the sadism subscale of the SD4 scale is the most reliable for measuring individuals with average to high levels ($\bar{x}$ \< $\Theta$ \< 2.5 *SD*) of sadism but suboptimal for measuring those with low levels of sadism [@blotnerDarkTriadDead2022]. Therefore, IRT models offer a more informative and nuanced perspective, proving useful for researchers who wish to closely examine the performance of their measures.

Further, IRT allows researchers to examine the performance of a specific item (hence "item" response theory) by specifying the relationship between item score and $\Theta$ given one-, two-, or three-parameter. One-parameter logistic (1PL) or Rasch model, the simplest variant of the IRT models, accounts only item difficulty (*b*) while presuming equal item discrimination (*a*) across items. Practically, this model slightly overlaps with CTT in the sense that they assume that all items carry equal informative value thus the estimated $\Theta$ of a 1PL/Rasch model is identical to the sum score of CTT [@langSciencePracticeItem2021; @stemlerRaschMeasurementItem2021]. This notion actually stems from a philosophical principle of *specific objectivity*, which requires that comparisons between measurements remain independent of both the item and individual characteristics [@raschSpecificObjectivityAttempt1977]. Put simply, Rasch model assumes that the latent trait ($\Theta$) should remain unaffected by specific items used in the test. For example, the difference in depression levels between two individuals should be always the same regardless of the scale used to measure their depression levels. While this principle reasonably enforces objectivity in measurement practices, it demands a strong theory delineating the construct and strict requirements of data-model fit while both assumptions rarely hold in a real-world scenario, especially in psychological science.

Furthermore, the two-parameter logistic (2PL) model fills this gap by allowing item ability to differentiate individuals with varying levels of $\Theta$ (i.e., item discrimination parameter - *a*) to differ. Additionally, in some contexts, researchers may suspect that a part of the probabilistic relationship between observed score and $\Theta$ is explained by guessing thus three-parameter logistic (3PL) model incorporates guessing parameter (*c*). The field of IRT has rapidly developed to include various models suited to specific contexts, such as handling ordinal responses [@murakiGeneralizedPartialCredit1992; @samejimaGradedResponseModel1997], categorical responses [@thissenNominalCategoriesItem2013], or assessing multidimensional traits simultaneously [@bockMarginalMaximumLikelihood1981; @chalmersMirtMultidimensionalItem2012]. It is important to note that 1PL, 2PL, and 3PL models are only applicable to binary or dichotomous data (e.g., true/false response), and the focus of this article is nonetheless to show the utility of IRT for fitting the ordered (Likert-style) responses. We briefly summarize the features of the most frequently employed IRT models in {apatb-table1}.

@@

The GRM is a family of IRT models specifically designed to analyze ordered polytomous (Likert-style) data [@samejimaGradedResponseModel1997; @samejimaGeneralGradedResponse2010; @samejimaGradedResponseModels2016]. The fundamental idea of GRM is to extend the logic of simpler dichotomous IRT models to polytomous scales by applying a probabilistic function to each response category. In dichotomous 2PL IRT models, the probability of answering a correct response is modeled as a function of $\Theta$, item difficulty, and item discrimination. GRM extends this idea by modeling the probability of picking a certain response category or higher on an item (i.e., *step function*) given to item difficulty and discrimination. For each response category (e.g., answering "agree" on a five-point Likert scale), GRM calculates the cumulative probability of a participant answering "agree" or above, given to their $\Theta$. Therefore, the notion of item difficulty in dichotomous IRT models is extended to the step function (i.e., item threshold) in order to handle ordered polytomous data.

To illustrate the step function, consider a scale measuring sadistic personality (e.g., "*watching a fist-fight excites me*," etc.) with response categories ranging from 1 (*strongly disagree*) to 4 (*strongly agree*). Imagine these categories are hierarchically ordered from the lowest category (*strongly disagree*) to the highest (*strongly agree*), like a staircase. GRM aims to estimate the level of sadistic personality ($\Theta$) required for "stepping" from one response category to another (e.g., *moderately agree* to *strongly agree*). Therefore, a GRM model of a five-point Likert scale calculates four item threshold parameters (*b*): the location of $\Theta$ level where individuals are equally likely to respond 1 or 2 (*b*~1~), 2 or 3 (*b*~2~), 3 or 4 (*b*~3~), and 4 or 5 (*b*~4~). Each threshold is exactly the point where a participant is equally likely to respond to either of the adjacent response categories. For example, if *b*~1~ = -1.23 for the item "*watching a fist-fight excites me*", it implies that participants with sadism level 1.23 below the mean are equally likely to either answer *strongly disagree* or *moderately disagree*.

Let's take a closer look at {apafg-fig1}, which shows an example of an item probability function (IPF) from an item of a scale with five response categories. In general, IPF describes a relationship between $\Theta$ (x-axis) and the probability of endorsing a response category (y-axis - *P*$\Theta$). {apafg-fig1} consists of five category probability curves, each of which represents the probability of endorsing a response category given $\Theta$. Item threshold is exactly the location of $\Theta$ where two adjacent category probability curves cross each other.

@@

Before running an analysis, researchers need to be aware of assumption underlying a GRM model. The first key assumption is that $\Theta$ is monotonically correlated with the probability of endorsing a response category. In other words, individuals with higher levels of $\Theta$ have a higher probability of endorsing response category that represents greater intensity of the traits [@hambletonIRTModelsAnalysis2010]. For instance, people with higher tendency of sadistic personality have a greater chance of answering "*moderately agree*" or "*strongly agree*" on a sadism scale compared to those with lower levels of sadism.

The second key assumption pertains to unidimensionality of $\Theta$. This implies that the item pool tested using a GRM analysis should represent only one latent trait ($\Theta$). Although it is indeed possible to account for multiple $\Theta$ in a IRT model (multidimensional IRT, see @bockFullInformationItemFactor1988, we limit the scope of our tutorial to the unidimensional GRM. An unidimensional IRT model can still be robustly applied to multidimensional data if multiple latent traits are moderately intercorrelated or when there is a strong general factor (*g*) underlying the data [@reiseEvaluatingImpactMultidimensionality2015]. To test whether an item pool represents only one $\Theta$, researchers can apply a parallel analysis [@guoAssessingDimensionalityIRT2023] or a factor analysis [@hambletonAssessingDimensionalitySet1986], before specifying their GRM models.

Relatedly, a GRM model assumes that the item pool are locally independent, as the third assumption. Local independence suggests that participants' responses to one item do not influence their response to another. It is important to note that evidence for unidimensionality does not always warrant local independence. As an eyeball example - consider a trust in science scale containing these two items: "*I trust scientists working in the natural science*" and "*most biologists are trustworthy*." While both items might measure the same latent trait (i.e., trust in science), those who trust scientists working in the natural science are very likely to also trust biologists. To inspect whether items are locally independent, researchers can examine the relationships between item responses after accounting for $\Theta$ using residual correlation analysis [@chenLocalDependenceIndexes1997].

@@

## Step 1: Preparation

@@

```{r}
install.packages("tidyverse", "psych", "devtools", "mirt", "caret", dependencies=TRUE)
devtools::install_github("masurp/ggmirt") # remote installation through GitHub repository
```

The above command also automatically requests to install dependency packages. Note that we install `ggmirt` directly from its GitHub repository because `ggmirt` is not yet available in the Comprehensive R Archive Network (CRAN). Next, we should activate the R packages by typing a command as follows:

@@

```{r}
zero <- colSums(rwa == 0) / nrow(rwa) * 100 # Computing the percentage of "0" in each column.
print(zero) # Examining the proportion of "0" for each item.
```

@@

After calling the commands, readers can find `Proportion Var` in the output, which reflects the amount of variance explained by a latent factor. The variance proportion of our EFA model is 0.56, which means that 56% of total variance is explained by a latent factor. Since the latent factor can account for more than 50% of variance, we can reasonably assume that the latent factor underlying our data is indeed unidimensional. To better visualize the eigenvalues of the EFA, we can draw a scree plot using these following commands:

@@

The first line of code is a function to plot the Eigenvalues (`efa$values`, x-axis) given the number of latent factors (y-axis). The option `type = "b"` implies that we ask R to display the Eigenvalues exact points and a line connecting these points along the number of factors. The second line of code aims to display a horizontal red line (i.e., a reference line) to see which factors have eigenvalues greater than one, which is a common criterion for factor retention. {apafg-fig2} shows that there is a clear break or an "elbow" that displays the location where the Eigenvalues start to level off. This implies that factors beyond 1 may not be meaningful and could just represent noise. Therefore, the scree plot strengthens our assumption that the item pool is, indeed, unidimensional.

To support unidimensionality assumption, we can also run a parallel analysis by running this following command:

@@

Readers may find a plot similar to {apafg-fig3} should appear right after executing this code, as well as a message, which says: `Parallel analysis suggests that the number of factors =  6  and the number of components =  NA`. This means that the parallel analysis suggests that there are six factors underlying our item pool, not one. However, according to {apafg-fig3}, we can see clearly that the Eigenvalues from our actual data fall steeply after the first factor. To paint a clearer picture, we can examine the Eigenvalues of each latent factor estimated by the parallel analysis by running this following command:

@@

@@

```{r}
fit <- mirt(data=rwa, 1, model=model, itemtype="graded", SE=T, verbose=F)
```

llalalala

```{r}
coefs <- coef(fit, IRTpars=T, simplify=T) # Storing model parameters in a data frame.
print(coefs) # Yielding model parameters: item discrimination (a) and threshold (b).
```

```{r}
summary(fit) 
```

bla bla lhalalala

```{r}
M2(fit, type="C2")
```

@@

According to our analysis, the model does not fit the data well (*M*(`r df`) = `r m2_stat`, *p* = `r p_value`, *RMSEA* = `r rmsea`, )

```{r}
item.fit <- itemfit(fit)
```

@@

```{r apatb-table4}
#| output: true
#| echo: false
#| apa-cap: Item Parameters
#| apa-note: λ = Standardized Factor Loadings, h2 = Commonality, a = Discrimination, b1-b8 = Response specific threshold parameters.
#| ft.align: left
#| column: page

>>>>

To scrutinize their measurement quality, psychological researchers have relied upon two foundational principles underlying psychological testing, i.e., CTT and IRT, to decompose observed score into its *deterministic* and *random* or stochastic elements [@zumboValidityFoundationalIssues2006]. On the one hand, CTT decomposes observed score into *true* score and its *error* components. In practice, researchers implement this principle by conducting Structural Equation Modeling (SEM), Exploratory Factor Analysis (EFA), or Confirmatory Factor Analysis (CFA). Specifying the measurement part of SEM, or employing EFA and CFA models, allows for the decomposition of the covariance matrix of a given data into latent factors, which represent true score or actual ability or trait level, and error terms (i.e., residuals or unique variances). Consequently, the relationship between observed score and true score is *linear*, as true score is essentially a linear transformation of observed score [@embretsonItemResponseTheory2000]. IRT, on the other hand, models this relationship in a *probabilistic* rather than a linear fashion. More sharply, an IRT model assumes a probabilistic relationship between the observed score and the latent trait being measured ($\theta$), and in particular, this probability takes into account the properties of the items, such as item discrimination, difficulty, and, if necessary, guessing [@bakerBasicsItemResponse2017].

@@

IRT offers an elegant solution to this issue by enabling the calculation of standard errors for each individual (*SE*~$\theta$~), and test reliability can be inferred from the average of *SE*~$\theta$~ across a sample of participants [@embretsonItemResponseTheory2000; @langSciencePracticeItem2021]. This approach allows for the estimation of reliability at varying levels of trait or ability (i.e., test information function - TIF). By doing this, when a test is overly challenging for low-performing individuals or too easy for high-performing individuals, IRT analysis can help identify the levels of trait or ability where the test is most reliable. A concrete example of this is a study scrutinizing the reliability of the Short Dark Tetrad (SD4) scale from an IRT perspective [@blotnerDarkTriadDead2022]. According to this study, the sadism subscale of the SD4 scale is the most reliable for measuring individuals with average to high levels ($\bar{x}$ \< $\theta$ \< 2.5 *SD*) of sadism but suboptimal for measuring those with low levels of sadism [@blotnerDarkTriadDead2022]. Therefore, IRT models offer a more informative and nuanced perspective, proving useful for researchers who wish to closely examine the performance of their measures.

Further, IRT allows researchers to examine the performance of a specific item (hence "item" response theory) by specifying the relationship between item score and $\theta$ given one-, two-, or three-parameter. One-parameter logistic (1PL) or Rasch model, the simplest variant of the IRT models, accounts only item difficulty (*b*) while presuming equal item discrimination (*a*) across items. Practically, this model slightly overlaps with CTT in the sense that they assume that all items carry equal informative value thus the estimated $\theta$ of a 1PL/Rasch model is identical to the sum score of CTT [@langSciencePracticeItem2021; @stemlerRaschMeasurementItem2021]. This notion actually stems from a philosophical principle of *specific objectivity*, which requires that comparisons between measurements remain independent of both the item and individual characteristics [@raschSpecificObjectivityAttempt1977]. Put simply, Rasch model assumes that the latent trait ($\theta$) should remain unaffected by specific items used in the test. For example, the difference in depression levels between two individuals should be always the same regardless of the scale used to measure their depression levels. While this principle reasonably enforces objectivity in measurement practices, it demands a strong theory delineating the construct and strict requirements of data-model fit while both assumptions rarely hold in a real-world scenario, especially in psychological science.

Furthermore, the two-parameter logistic (2PL) model fills this gap by allowing item ability to differentiate individuals with varying levels of $\theta$ (i.e., item discrimination parameter - *a*) to differ. Additionally, in some contexts, researchers may suspect that a part of the probabilistic relationship between observed score and $\theta$ is explained by guessing thus three-parameter logistic (3PL) model incorporates guessing parameter (*c*). The field of IRT has rapidly developed to include various models suited to specific contexts, such as handling ordinal responses [@murakiGeneralizedPartialCredit1992; @samejimaGradedResponseModel1997], categorical responses [@thissenNominalCategoriesItem2013], or assessing multidimensional traits simultaneously [@bockMarginalMaximumLikelihood1981; @chalmersMirtMultidimensionalItem2012]. It is important to note that 1PL, 2PL, and 3PL models are only applicable to binary or dichotomous data (e.g., true/false response), and the focus of this article is nonetheless to show the utility of IRT for fitting the ordered (Likert-style) responses. We briefly summarize the features of the most frequently employed IRT models in {apatb-table1}.

@@

The GRM is a family of IRT models specifically designed to analyze ordered polytomous (Likert-style) data [@samejimaGradedResponseModel1997; @samejimaGeneralGradedResponse2010; @samejimaGradedResponseModels2016]. The fundamental idea of GRM is to extend the logic of simpler dichotomous IRT models to polytomous scales by applying a probabilistic function to each response category. In dichotomous 2PL IRT models, the probability of answering a correct response is modeled as a function of $\theta$, item difficulty, and item discrimination. GRM extends this idea by modeling the probability of picking a certain response category or higher on an item (i.e., *step function*) given to item difficulty and discrimination. For each response category (e.g., answering "agree" on a five-point Likert scale), GRM calculates the cumulative probability of a participant answering "agree" or above, given to their $\theta$. Therefore, the notion of item difficulty in dichotomous IRT models is extended to the step function (i.e., item threshold) in order to handle ordered polytomous data.

To illustrate the step function, consider a scale measuring sadistic personality (e.g., "*watching a fist-fight excites me*," etc.) with response categories ranging from 1 (*strongly disagree*) to 4 (*strongly agree*). Imagine these categories are hierarchically ordered from the lowest category (*strongly disagree*) to the highest (*strongly agree*), like a staircase. GRM aims to estimate the level of sadistic personality ($\theta$) required for "stepping" from one response category to another (e.g., *moderately agree* to *strongly agree*). Therefore, a GRM model of a five-point Likert scale calculates four item threshold parameters (*b*): the location of $\theta$ level where individuals are equally likely to respond 1 or 2 (*b*~1~), 2 or 3 (*b*~2~), 3 or 4 (*b*~3~), and 4 or 5 (*b*~4~). Each threshold is exactly the point where a participant is equally likely to respond to either of the adjacent response categories. For example, if *b*~1~ = -1.23 for the item "*watching a fist-fight excites me*", it implies that participants with sadism level 1.23 below the mean are equally likely to either answer *strongly disagree* or *moderately disagree*.

Let's take a closer look at {apafg-fig1}, which shows an example of an item probability function (IPF) from an item of a scale with five response categories. In general, IPF describes a relationship between $\theta$ (x-axis) and the probability of endorsing a response category (y-axis - *P*$\theta$). {apafg-fig1} consists of five category probability curves, each of which represents the probability of endorsing a response category given $\theta$. Item threshold is exactly the location of $\theta$ where two adjacent category probability curves cross each other.

@@

Before running an analysis, researchers need to be aware of assumption underlying a GRM model. The first key assumption is that $\theta$ is monotonically correlated with the probability of endorsing a response category. In other words, individuals with higher levels of $\theta$ have a higher probability of endorsing response category that represents greater intensity of the traits [@hambletonIRTModelsAnalysis2010]. For instance, people with higher tendency of sadistic personality have a greater chance of answering "*moderately agree*" or "*strongly agree*" on a sadism scale compared to those with lower levels of sadism.

The second key assumption pertains to unidimensionality of $\theta$. This implies that the item pool tested using a GRM analysis should represent only one latent trait ($\theta$). Although it is indeed possible to account for multiple $\theta$ in a IRT model (multidimensional IRT, see @bockFullInformationItemFactor1988, we limit the scope of our tutorial to the unidimensional GRM. An unidimensional IRT model can still be robustly applied to multidimensional data if multiple latent traits are moderately intercorrelated or when there is a strong general factor (*g*) underlying the data [@reiseEvaluatingImpactMultidimensionality2015]. To test whether an item pool represents only one $\theta$, researchers can apply a parallel analysis [@guoAssessingDimensionalityIRT2023] or a factor analysis [@hambletonAssessingDimensionalitySet1986], before specifying their GRM models.

Relatedly, a GRM model assumes that the item pool are locally independent, as the third assumption. Local independence suggests that participants' responses to one item do not influence their response to another. It is important to note that evidence for unidimensionality does not always warrant local independence. As an eyeball example - consider a trust in science scale containing these two items: "*I trust scientists working in the natural science*" and "*most biologists are trustworthy*." While both items might measure the same latent trait (i.e., trust in science), those who trust scientists working in the natural science are very likely to also trust biologists. To inspect whether items are locally independent, researchers can examine the relationships between item responses after accounting for $\theta$ using residual correlation analysis [@chenLocalDependenceIndexes1997].

@@

## Step 1: Data Preparation

@@

```{r}
install.packages("tidyverse", # for data wrangling
                 "psych", # for descriptive statistics and unidimensionality test
                 "devtools", # for installing R package that is not available on CRAN
                 "mirt", # for conducting the main GRM analysis
                 "caret", # for helping us detecting large residuals correlation (local independence test)
                 dependencies=TRUE)
devtools::install_github("masurp/ggmirt") # remote ggmirt installation through GitHub repository
```

The above command also automatically requests to install dependency packages. Note that we install *ggmirt* directly from its GitHub repository because *ggmirt* is not yet available in the Comprehensive R Archive Network (CRAN). Next, we should activate the R packages by typing a command as follows:

@@

```{r}
zero <- colSums(rwa == 0) / nrow(rwa) * 100 # Computing the percentage of "0" in each column.
print(zero) # Showing the percentage of "0" for each item.
```

@@

After calling the commands, readers can find `Proportion Var` in the output, which reflects the amount of variance explained by a latent factor. The variance proportion of our EFA model is 0.56, which means that 56% of total variance is explained by a latent factor. Since the latent factor can account for more than 50% of variance, we can reasonably assume that the latent factor underlying our data is unidimensional. To better visualize the Eigenvalues of the EFA, we can draw a scree plot using these following commands:

@@

The first line of code is a function to plot the Eigenvalues (`efa$values`, x-axis) given the number of latent factors (y-axis). The option `type = "b"` implies that we ask R to display the Eigenvalues exact points and a line connecting these points along the number of factors. The second line of code aims to display a horizontal red line (i.e., a reference line) to see which factors have eigenvalues greater than one, which is a common criterion for factor retention. {apafg-fig2} shows that there is a clear break or an "elbow" that displays the location where the Eigenvalues start to level off. This implies that factors beyond the first may not be meaningful and could be only noise. Therefore, the scree plot strengthens our assumption that the item pool is, indeed, unidimensional.

To scrutinize unidimensionality assumption, we can also run a parallel analysis by running this following command:

@@

Readers should find a plot similar to {apafg-fig3} appear right after executing this code as well as a message, which says: `Parallel analysis suggests that the number of factors =  6  and the number of components =  NA`. This means that the parallel analysis suggests that there are six factors underlying our item pool, not one. However, according to {apafg-fig3}, we can see clearly that the Eigenvalues from our actual data fall steeply after the first factor. To paint a clearer picture, we can examine the Eigenvalues of each latent factor estimated by the parallel analysis by running this following command:

@@

In this step, we start to specify a vector representing our model specification, such as:

@@

This code implies that we want to estimate a model with $\theta$ namely `rwa`, and the items representing the $\theta$ are column 1 to 22 of the data frame that we are going to use in the analysis. Then, we can continue to model fitting by typing this following command:

```{r}
fit <- mirt(data = rwa, 1, model = model, itemtype = "graded", SE = TRUE, verbose = FALSE)
```

The code above implies that we are fitting an IRT model using data from a data frame namely `rwa` (`data = rwa`). The number `1` indicates that we presume that the model contains only one $\theta$, i.e., unidimensional. We also set the model specification to `model = model`, which we have previously determined. We set `itemtype` to `"graded"` because we want to run a GRM model. The `SE` option is set to `TRUE` because we ask R to also estimate the standard error. At last, we set the `verbose` option to `FALSE` to avoid R displaying unnecessary information during the model fitting process, so the `mirt()` function can run quietly.

Apart from this tutorial, readers might encounter four common error messages after running `mirt()` function: (a) `EM cycles terminated after 500 iterations`, (b) `Could not invert information matrix; model may not be empirically identified`, (c) `The following items have only one response category and cannot be estimated`, and (d) `Error in [<-(*tmp*, i, nms[[i]], value = items.old[[i]]) : subscript out of bounds`. Error messages (a) and (b) correspond to the convergence issue, while (c) and (d) are issues related to the data structure.

*mirt* uses the EM algorithm as a default option. In short, the EM algorithm estimates model parameters in two steps: the expectation (E) and the maximization (M) steps. In the E step, the algorithm estimates the expected value of the $\theta$, and in the M step, it improves model parameters to maximize the likelihood of the observed data given the expected value of $\theta$. The process continues until model parameters are converged in stable values that fit to the data. If the error message (a) appears, it means that the package cannot estimate a stable solution for model parameters after 500 iterations. As a solution, readers can simply increase the maximum number of iterations to, for instance, 1000 cycles by adding this option `technical = list(NCYCLES = 10000)` to the `mirt()` function.

The solution to error message (b) is less straightforward than (a) because it indicates that the model cannot be identified at all, which can have several causes. It could be due to an error in specifying assumptions about the number of latent traits or item properties, or it could be due to insufficient sample size. Therefore, when error message (b) appears, it is important to take a step back and think twice about the model assumptions, the fit between the model and the underlying theory, and whether the sample size is large enough to estimate the model properly.

The error message (c) and (d) are easier to solve because both indicate the error are related to the data structure. The former means that certain items have no variance, consisting only one response category. The solution for this issue is to exclude these items from model estimations. The latter indicates a mismatch between the vector of model specification and the data frame, for instance, specifying a vector with ten variables, while in fact, the data frame has only nine columns. To solve this issue, readers are suggested to check again whether the specified vector representing the model specification matches the structure of the data frame.

Next, we can ask the package to calculate item discrimination and thresholds by calling these functions below:

```{r}
coefs <- coef(fit, IRTpars=TRUE, simplify=TRUE) # Saving model parameters (item discrimination (a) and threshold (b)) in a data frame.
print(coefs) # Calling the data frame.
summary(fit) # Displaying factor loadings and commonality.
```

The first line of code indicates that we are calling a `coef()` function to calculate item parameters from `fit` model, and then, store them in a new data frame called `coefs`. The `IRTpars` and `simplify` option is set to `TRUE` because we want to display the traditional IRT parameters (*a* and *b*) in a simple, readable format without showing the standard errors.

The output from this procedure is presented in {apatb-table4}, which shows item discrimination (*a*) of each item and item thresholds (*b1-b8*) of each response category within an item. To interpret *a*, readers can use a rule of thumb suggested by @bakerBasicsItemResponse2017, which categorize item discrimination exceeding 0.00, 0.35, 0.65, 1.35, and 1.70, as very low, low, moderate, high, and very high. As we seen in {apatb-table4}, item discrimination parameters of the RWA scale are ranging from high to very high (1.57 (Q1) to 3.32 (Q7)), suggesting efficient differentiation among individuals at different RWA levels. All threshold parameters  show a consistent increase, implying that those with higher RWA levels tend towards higher response categories. However, *b1* and *b2* for all items are very close to zero, which indicates that individuals with RWA level around the mean are likely to opt for "*strongly disagree*" or stepping to its adjacent response category. This might suggest that the RWA scale is less sensitive to measure individuals with a low or very low RWA level.

The third line of code is a function to calculate a factor loading ($\lambda$) and commonality (h^2^) of each item, which both are also provided in {apatb-table4}. Before estimating a graded response model, *mirt* ran an EFA, and now as we can see in the console, we are looking at the EFA results. The results are slightly different from our previous EFA analysis, because *mirt* runs EFA using a quasi polychoric correlation matrix, while the one we ran earlier to test unidimensionality used a Pearson's correlation matrix as an input. However, most importantly, we see that all items are significantly loaded to one factor, and the factor now substantially accounts for 65.1% of the variance in the data, which strengthens our assumption that the RWA scale is unidimensional.

Next, we can estimate the model goodness-of-fit (GOF) statistics (*M*~2~) by calling this command below:

```{r}
M2(fit, type="C2")
```
```{r}
#| include: false

@@

We ask the package to calculate the scaled *M*~2~ statistics [@maydeu-olivaresLimitedInformationGoodnessoffit2006] and other GOF indices, such as Root Mean Square Error of Approximation (RMSEA), Comparative Fit Index (CFI), Tucker-Lewis Index (TLI), and Standardized Root-Mean-Square Residual (SRMSR) by running this code above. Here, we set the `type` option to `"C2"` because it is suitable for computing *M*~2~ statistics in polytomous models. The interpretation of GOF statistics are similar to a general guide of interpreting GOF in CFA or SEM models [@schermelleh-engelEvaluatingFitStructural2003].

According to our analysis, the model overall does not fit the data well (*M*~2~(`r df`) = `r m2_stat`, *p* = `r p_value`, *RMSEA* = `r rmsea`, *SRMSR* = `r srmsr`, TLI = `r tli`, CFI = `r cfi`), and this may be caused by local dependency between the items. We will explore this issue further in the next part.

Next, we calculate item-level fit statistics, i.e., the scaled $\chi$^2^ statistics, by running this following function:

```{r}
item.fit <- itemfit(fit) # Estimating item fit statistics and saving them in a data frame.
print(item.fit) # Calling the data frame.
```

The output of this function is provided in {apatb-table5}, which shows fit statistics for an individual item. All items have very good (low) *RMSEA* values but normally, we want the *p* values of the scaled $\chi$^2^ to be nonsignificant, implying there is no discrepancy between the model and data. In {apatb-table5}, only three items here are nonsignificant (Q2, Q4, and Q12), which strengthens our suspicion that the RWA items might not be locally independent.

@@

```{r apatb-table4}
#| output: true
#| echo: false
#| apa-cap: Item Parameters
#| apa-note: a = Discrimination, b1-b8 = Response specific threshold parameters, λ = Standardized Factor Loadings, h2 = Commonality.
#| ft.align: left
#| column: page
